AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-14 Paper/main.tex
10-14 paper

 \citeauthor{MuhammadRafid2020} consider deep-learning-based models as black boxs and believe they lack interpretability, 

@@

Attention mechanism has demonstrated its power in Natural Language Processing, Statistical Learning, Speech and Computer Vision \citep{chaudhari2019an}. 
 It makes model tends to focus selectively on parts of the input which is help in performing the task effectively. 
 Previous observation have shown that Cas9 preferentially binds sgRNAs containing purines but not pyrimidines \citep{wang2014genetic} and multiple thymine in the spacer impairing sgRNA activity \citep{wu2014genome-wide}, 
 that is to say some specific nucleotide and base position need more attention compared to others. 
 The above is the premise of introducing attention mechanism. Strictly speaking, we are not the first to bring attention mechanisms into this field. 
 The most similar approach to ours is the work based on transformer by \citeauthor{Liu2019}. 
 They use transformer, a components based on attention mechanism, instead of RNN to improve the ability of temporal feature extraction, 
 hence, enhance the performance of their model \citep{Liu2019,vaswani2017attention}. 
 In our work,the interpretability benefit from attention mechanism is more focused.
 In this paper, our main contributions are as follows:\vspace*{1pt}
\begin{itemize}
    \item Present a novel deep-learning-based model, which can extract potential feature representation of sgRNA sequence in both spatial and temporal domain parallelly. 
    Finally, the ensemble learning method is used to combine the two to achieve better performance than other models.
    It doesn't belong to any of the above categories of existing approaches.
    \item Introduce attention mechanism into our model. 
    As a result, It does not need post hoc explanations techniques based on input perturbation to explain itself. 
    It is intrinsic interpretable in both temporal and spatial domains.
    In the spatial domain it's at the nucleotide level, in our word, first-order preference, 
    while at the overall level in the temporal domain, in our word, second-order preference. 
    Thus, it is transformed from a black box to an intrinsically interpretable model with the performance of deep learning based model. 
    \item Throught ablation analysis and testing a series of possible network structure, 
    we find there are multiple components and strategy can improve the performance of our model and constructed AttCRISPR, 
    which could outperform the current state-of-the-art tool on DeepHF dataset.\vspace*{1pt}
\end{itemize}

@@

\begin{figure}[!tpb]%figure3

@@

Two different sets of experiments are carried out in our work. 
The first one is designed for ablation analysis of AttCRISPR. 
We compare the performance of method in both spatial and temporal domain (without any hand-crafted biological features). 
Furthermore, we test the ensemble method based on the same strategy, which proved that the ensemble of method in both spatial and temporal domain can significantly improve the performance.

@@

Our baselines have a comprehensive coverage of the methods tested in these datasets.

In Table~\ref{Tab:01}, we annotate some properties of these baselines (with/without neural models, with/without hand-crafted feature)

\begin{table}[!t]
    \processtable{
    The naming convention code and brief description of the methods compared in our work\label{Tab:01}} 
    {\begin{tabular}{@{}lllll@{}}\toprule 
        Method & Code & Neural & Hand-crafted Feature & Description\\\midrule
        Linear$^*$ & row1 & No & No & row1\\
        Ridge$^*$ & row4 & No & No & row4\\
        XGBoost$^*$ & row1 & No & No & row1\\
        MLP$^*$ & row1 & Yes & No & row1\\
        DeepHF$^*$ & row1 & Yes & Yes & row1\\
        CRISPRpred(SEQ)$^{\#}$ & row1 & No & Yes & row1\\
        \botrule
    \end{tabular}}{This is a footnote}
\end{table}
\section{Results}

@@

\enlargethispage{6pt}

@@

\subsection{Test1}

>>>>

 \citeauthor{MuhammadRafid2020} consider deep-learning models as black boxs and believe they lack interpretability, 

@@

Attention mechanism has demonstrated its power in Natural Language Processing, Statistical Learning, Speech and Computer Vision \citep{chaudhari2019an}. 
 It makes model tends to focus selectively on parts of the input which is help in performing the task effectively. 
 Previous observation have shown that Cas9 preferentially binds sgRNAs containing purines but not pyrimidines \citep{wang2014genetic} and multiple thymine in the spacer impairing sgRNA activity \citep{wu2014genome-wide}, 
 that is to say some specific nucleotide and base position need more attention compared to others. 
 The above is the premise of introducing attention mechanism. Strictly speaking, we are not the first to bring attention mechanisms into this field. 
 The most similar approach to ours is the work based on transformer by \citeauthor{Liu2019}. 
 They use transformer, a components based on attention mechanism, instead of RNN to improve the ability of temporal feature extraction, 
 hence, enhance the performance of their model \citep{Liu2019,vaswani2017attention}. 
 In our work,the interpretability benefit from attention mechanism is more focused.
 In this paper, our main contributions are as follows:\vspace*{1pt}
\begin{itemize}
    \item Present a novel deep-learning model, which can extract potential feature representation of sgRNA sequence in both spatial and temporal domain parallelly. 
    Finally, the ensemble learning method is used to combine the two to achieve better performance than other models.
    It doesn't belong to any of the above categories of existing approaches.
    \item Introduce attention mechanism into our model. 
    As a result, It does not need post hoc explanations techniques based on input perturbation to explain itself. 
    It is intrinsic interpretable in both temporal and spatial domains.
    In the spatial domain it's at the nucleotide level, in our word, first-order preference, 
    while at the overall level in the temporal domain, in our word, second-order preference. 
    Thus, it is transformed from a black box to an intrinsically interpretable model with the performance of deep learning based model. 
    \item Throught ablation analysis and testing a series of possible network structure, 
    we find there are multiple components and strategy can improve the performance of our model and constructed AttCRISPR, 
    which could outperform the current state-of-the-art tool on DeepHF dataset.\vspace*{1pt}
\end{itemize}

@@

@@

Two different sets of experiments are carried out in our work. 
The first one is designed for ablation analysis of AttCRISPR. 
We compare the performance of end2end method (without any hand-crafted biological features) in both spatial and temporal domain. 
Furthermore, we test the ensemble method based on the same strategy, which proved that the ensemble of method in both spatial and temporal domain can significantly improve the performance.

@@

\begin{table}[!tpb]
    \processtable{The methods compared in our work and brief description of these methods\label{Tab:01}} 
    {\begin{tabular}{@{}lllp{3.5cm}l@{}}\toprule
        Method & Neural & End2end & Description\\\midrule
        CNN$^*$ & Yes & Yes & Naive convolutional neural network\\
        RNN$^*$ & Yes & Yes & Bidirectional long short-term memory neural network\\
        XGBoost$^*$ & No & Yes & Extreme Gradient Boosting regression tree\\
        MLP$^*$ & Yes & Yes & Multilayer perceptron\\
        DeepHF$^*$ & Yes & No & Bidirectional long short-term memory neural network (with hand-crafted biological features)\\
        CRISPRpred(SEQ)$^{\#}$ & No & No & A conventional machine learning pipeline\\
        SpAC & Yes & Yes & Spatial AttCRISPR\\
        TAC & Yes & Yes & Temporal AttCRISPR\\
        EnAC & Yes & Yes & Ensemble AttCRISPR (without hand-crafted biological features)\\
        StAC & Yes & No & Standard AttCRISPR\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
We use the conclusion of corresponding paper as the performance indicators of the model directly. 
Specially, CRISPRpred(SEQ) take another set of hand-crafted sequence-based feature to improve performance.}
\end{table}

Our baselines have a comprehensive coverage of the methods tested in these datasets. 
In Table~\ref{Tab:01}, we annotate some properties of these baselines (is/isn't neural models, is/isn't end2end models)
All the experiments were carried out in Python 3.6 using Keras 2.2.4 and one GeForce RTX 2080Ti Super was used for training and testing if needed.

\section{Results}

We designed experiments to address the following questions:
\begin{itemize}
    \item 
    Whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

\subsection{Model building and stacking}\label{section:stacking}

@@

\subsection{Comparison to current methods}\label{section:comparison}

@@

\subsection{Interpretability of the AttCRISPR}\label{section:interpretability}
